Predicting information integration in pragmatic word learning across development

### 2021-01-05 papers/full_paper/manuscript.Rmd
prep for re-submission

abstract: |
  Language is learned in complex social settings where listeners must reconstruct speakers’ intended meanings from context. To navigate this challenge, children can use pragmatic reasoning to learn the meaning of unfamiliar words. One important challenge for pragmatic reasoning is that it requires integrating multiple information sources. Here we study this integration process. We isolate two sources of pragmatic information (common ground and expectations about informativeness) and -- using a probabilistic model of conversational reasoning -- formalize how they should be combined and how this process might develop. We use this model to generate quantitative predictions, which we test against new behavioral data from three- to five-year-old children (N = 243) and adults (N = 694). Results show close numerical alignment between model predictions and data. Furthermore, the model provided a better explanation of the data compared to simpler alternative models assuming that children selectively ignore one information source. This work integrates distinct sets of findings regarding early language and suggests that pragmatic reasoning models can provide a quantitative framework for understanding developmental changes in language learning.

@@

Children learning their first language make inferences about intended meanings based on utterance-level and common-ground information both for language understanding and language learning [@bohn2019pervasive; @clark2009first; @tomasello2008origins]. Starting very early, infants expect adults to produce utterances in a cooperative way [@behne2005one], and expect language to be carrying information [@vouloumanos2012twelve]. By age two, children are sensitive to the informativeness of communication [@o2001two]. By age three children can use this expectation to make pragmatic inferences [@stiller2015ad; @yoon2019role] and to infer novel word meanings [@frank2014inferring]. And although older children continue to struggle with some complex pragmatic inferences until age five and beyond [@noveck2001children], an emerging consensus identifies these difficulties as stemming from difficulties reasoning about linguistic alternatives rather than pragmatic deficits [@skordos2016children; @horowitz2018trouble; @barner2011accessing]. Thus, children’s ability to reason about utterance-level pragmatics is present at least by ages three to five, and possibly substantially younger.

Common ground has traditionally been defined in recursive terms: in order to be part of common ground, some piece of information has to be not just known to both interlocutors but also known to both to be shared between them [@clark1996using]. Numerous studies probed the role of sharedness of information and found that it plays a critical role in communicative interactions [@brown2009partner; @hanna2003effects; @heller2016perspective; @mozuraitis2015privileged]. Based on this literature, one might argue that the term common ground should be restricted to describe situations in which the sharedness aspect is directly tested. However, most of this work is focused on online perspective taking. In this paper, we use the term common ground to refer to shared information that is built up over the course of an interaction - something that is likely easier for children [@matthews2006effect]. We assume that the consequence of a direct interaction (with matching perspectives) between the speaker and the listener is that information is mutually manifest; that is, not just known to both interlocutors but also assumed to be shared between them [@bohn2018common] and hence part of common ground. Thus, since this information is unproblematically in common ground, we can focus on how this information integrates with other pragmatic information sources.

Evidence for the use of common ground information by young children is even stronger: Common ground information guides how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social; @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role; @bohn_le_peloquin_koymen_frank_2020; @saylor2009preschoolers; @sullivan2019discourse].

@@

RSA models capture common ground information as a shared prior distribution over possible intended meanings. Thus, a natural locus for information integration within probabilistic models of pragmatic reasoning is the trade off between the prior probability of a meaning and the informativeness of the utterance. This trade off between contextual factors during word learning is a unique aspect that is not addressed by other computational models of word learning, which have focused on learning from cross-situational, co-occurrence statistics [@fazly2010probabilistic; @frank2009using] or describing generalizations about word meaning [@xu2007word].

@@

We start by replicating previous findings with adults showing that listeners make pragmatic inferences based on non-linguistic properties of utterances in isolation (experiment 1). Then we show that adults make inferences based on common ground information (experiment 2A and 2B). We use data from these experiments as parameters to generate a priori predictions from RSA models about how utterance and common ground information should be integrated. We consider three models that make different assumptions about the integration process: In the *integration model*, the two information sources are integrated with one another. The other two models are lesion models that assume that participants focus on one type of information and disregard the other whenever they are presented together. According to the *no common ground* model, participants focus only on the utterance information and in the *no informativeness* model, only common ground information is considered. We compare predictions from these models to new empirical data from experiments in which utterance and common ground information are manipulated simultaneously (Experiment 3 and 4).

@@

Taken together, this work makes two primary contributions: first, it shows that both adults and children integrate utterance-level and common-ground information flexibly. Second, it uses Bayesian data analysis within the RSA framework to provide a model for understanding the multiple loci for developmental change in complex behaviors like contextual communication.

@@

We preregistered sample sizes, inferential statistical analysis and computational models for all experiments. All deviations from the registered analysis plan are explicitly mentioned. All analyses were run in R [@R-base]. All p-values are based on two sided analysis. Cohen's d (computed via the function `cohensD`) was used as effect size for t-tests. Frequentist logistic GLMMs were fit via the function `glmer` from the package `lme4` [@R-lme4] and had a maximal random effect structure conditional on model convergence. Details about GLMMs including model formulas for each experiment can be found in the Supplementary Material.

All models and model comparisons were implemented in WebPPL [@dippl] using the R package `rwebppl` [@R-rwebppl]. Probabilistic models were evaluated using Bayesian data analysis [@lee2014bayesian], also implemented in WebPPL. In experiment 3, 4 and 7, we compared probabilistic models based on Bayes Factors - the ratio of the marginal likelihoods of each model given the data. Details on models, including information about priors for parameter estimation and Markov chain Monte Carlo settings can be found in the Supplementary Material available online. Code to run the models is available in the associated online repository.

@@

Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model. We modeled the integration of utterance informativity and common ground as a process of socially-guided probabilistic inference, using the results of experiments 1 and 2 to inform key parameters of a computational model. The Rational Speech Act (RSA) model architecture introduced by Frank and Goodman [-@frank2012predicting] encodes conversational reasoning through the perspective of a listener (“he” pronoun) who is trying to decide on the intended meaning of the utterance he heard from the speaker (“she” pronoun). The basic idea is that the listener combines his uncertainty about the speaker’s intended meaning - a prior distribution over referents P(r) - with his generative model of how the utterance was produced: a speaker trying to convey information to him. To adapt this model to the word learning context, we enrich this basic architecture with a mechanism for expressing uncertainty about the meanings of words (lexical uncertainty) - a prior distribution over lexica P(L) [@bergen2016pragmatic].

@@

In the above equation, the listener is trying to jointly resolve the speaker’s intended referent r and the meaning of words (thus learning the lexicon $\mathcal{L}$). He does this by imagining what a rational speaker would say, given the referent they are trying to communicate and a lexicon. The speaker is an approximately rational Bayesian actor (with degree of rationality alpha), who produces utterances as a function of their informativity. The space of utterances the speaker could produce depends upon the lexicon $P(u|\mathcal{L})$; simply put, the speaker labels objects with the true labels under a given lexicon L (see Supplementary Material available online for details):

@@

This computational model provides a natural avenue to formalize quantitatively how informativeness and common ground trade-off during word learning. As mentioned above, the common ground shared between speaker and listener plays the role of the listener’s prior distribution over meanings, or types of referents, that the speaker might be referring to and which we posit depends on prior interactions around the referents in the present context (e.g., preference or novelty; experiment 2A and B). We use the results from experiment 2 to specify this distribution. The in-the-moment, contextual informativeness of the utterance is captured in the likelihood term, whose value depends on the rationality parameter $\alpha$. Assumptions about rationality may change depending on context and we therefore used the data from experiment 1 to specify $\alpha$ (see Supplementary Material available online for details about these parameters).

@@

As a second step, we compared the model predictions to the data. Participants’ average responses were highly correlated with the  predictions from the *integration model* in each condition (Fig \ref{fig:fig3}b). When comparing model, we found that model fit was considerably better for the *integration model* compared to the *no common ground model* (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information.

Finally, we examined the noise parameter for each model. The estimated proportion of random responses according to the *integration model* was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This parameter was substantially lower for the *integration model* compared to the alternative models (*no common ground model*: `r ex3_flat_prior_model_noise_param%>%pull(mean)` [`r ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex3_prior_only_model_noise_param%>%pull(mean)` [`r ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex3_prior_only_model_noise_param%>%pull(ci_upper)`]), lending additional support to the conclusion that the *integration model* better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

@@

To test if the *integration model* makes accurate predictions for different combinations, we first replicated and then extended the results of experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (3 levels - strong, medium and weak - for preference and 2 levels - strong and medium - for novelty) by changing the way the speaker interacted with the objects prior to the request. The procedural details and statistical analysis for these these manipulations are described in the Supplementary Material available online. For experiment 4, we paired each level of prior strength manipulation with the informativeness inference in the same way as in experiment 3. This resulted in a total of 20 conditions, for which we generated a priori model predictions in the same way as in experiment 3. That is, we conducted a separate experiment for each level of prior strength and common ground manipulation to estimate the prior probability of each object following this particular manipulation (analogous to experiment 2). This prior distribution was then passed through the model for the congruent and incongruent conditions, resulting in a unique prediction for each of the 20 condition. Given the graded nature of the prior manipulations, experiment 4 basically tests how well the model performs with different types of prior distributions.

@@

The direct replication of experiment 3 within experiment 4 showed a very close correspondence between the two rounds of data collection (see Fig \ref{fig:fig3}c). GLMM results for experiment 4 can be found in the Supplementary Material available online. Here we focus on the analysis based on the probabilistic models. Model predictions from the *integration model* were again highly correlated with the average response in each condition (see Fig \ref{fig:fig3}d). We evaluated model fit for the same models as in experiment 3 and found again that the *integration model* fit the data much better compared to the *no common ground* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the *integration model* was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to experiment 3 and again lower compared to the alternative models (*no common ground model*: `r ex4_flat_prior_model_noise_param%>%pull(mean)` [`r ex4_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex4_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex4_prior_only_model_noise_param%>%pull(mean)` [`r ex4_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex4_prior_only_model_noise_param%>%pull(ci_upper)`]).

@@

In experiment 6, we assessed whether children use common ground information to identify the referent of a novel word. We tested children only with the preference manipulation^[We initially tested children with the novelty as well as the preference manipulation. We found that children made the basic inference in that they selected the object that was preferred by or new to the speaker, but found little evidence that children distinguished between requests made by the same speaker or a different speaker in the case of novelty. This finding contrasts with earlier work [@diesendruck2004two]. Since our focus was on how children integrate informativeness and common ground, we did not follow up on this finding but dropped the novelty manipulation and focused on preference for the remainder of the study.]. The procedure for children was identical to the preference manipulation for adults. Children received eight trials, four with the same and four with a different speaker.

@@

child_ex3_bf <- readRDS("../../stats/saves/child_ex3_model_comparison.rds") %>%
  group_by(parameter)%>%
  spread(model, logP) %>%
  mutate("pragmatic_flat_prior" = exp(pragmatic - flat_prior),
         "pragmatic_prior_only" = exp(pragmatic - prior_only),
         "flat_prior_prior_only" = exp(flat_prior - prior_only)) %>%
  select(-pragmatic,-flat_prior, -prior_only)

@@

Analyses comparing the model predictions from the probabilistic models to the data suggest that children flexibly integrate both common ground and informativity information. Furthermore, this integration process is accurately captured by the *integration model* at least for four-year-olds. For the correlational analysis, we binned model predictions and data by year. There was a substantial correlation between the predicted and measured average response for four-year-olds, but less so for three-year-olds (Fig \ref{fig:fig5}b). One of the reasons for the latter was the low variation between conditions. For the model comparison, we treated age continuously. As with adults, we found a much better model fit for the *integration model* compared to the *no common ground* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_flat_prior), digits = 2)`) or the *no informativeness model* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_prior_only), digits = 2)`).

The inferred level of noise based on the data for the integration model was `r child_ex3_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r child_ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r child_ex3_pragm_model_noise_param%>%pull(ci_upper)`), which was lower compared to the alternative models considered (*no common ground model*: `r child_ex3_flat_prior_model_noise_param%>%pull(mean)` [`r child_ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r child_ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r child_ex3_prior_only_model_noise_param%>%pull(mean)` [`r child_ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r child_ex3_prior_only_model_noise_param%>%pull(ci_upper)`]) but numerically higher than that of adults.

The high level of inferred noise moved the model predictions for children in all conditions close to chance level. We therefore compared two additional sets of models with different parameterizations of the noise parameter that emphasized differences between conditions in the model predictions more (see Supplementary Material and Fig \ref{fig:fig5}a). This analysis was not pre-registered. Parameter free models did not include a noise parameter and developmental noise models allowed the noise parameter to change with age. In each case, the *integration model* provided a better fit compared to the alternative models (*no common ground*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_flat_prior), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_flat_prior), digits = 2)`; *no informativeness*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_prior_only), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_prior_only), digits = 2)`). The developmental noise parameter for the integration model decreased with age, suggesting that older children behaved more in line with model predictions compared to younger children (see Fig. S13 in Supplementary Material available online).

```{r fig5, include = T, fig.align = "center", fig.cap = "Results from experiment 7 for children. Model predictions and data across age in the four conditions (a). Transparent black dots show data from individual participants and black lines show conditional means of the data with 95\\% CI. Black diamonds show the mean of the data for age bins by year and error bars show 95\\% CIs. Correlation between model predictions (with noise parameter) and condition means binned by year (b). Coefficients and p-values are based on Pearson correlation statistic. Error bars and shaded regions represent 95\\% HDIs. For 4-year-olds, two conditions yielded the same data means and model predicitons and are thus plotted on top of each other.", out.width = "80%"}
knitr::include_graphics("./figures/fig5.png")
```

@@

Integrating multiple sources of information is an integral part of human communication. To infer the intended meaning of an utterance, listeners must combine their knowledge of communicative conventions (semantics and syntax) with social expectations about their interlocutor. This integration is especially vital in early language learning, and the different varieties of pragmatic information are among the most important sources [@bohn2019pervasive]. But how are pragmatic cues integrated during word learning? Here we used a Bayesian cognitive model to formalize this integration process. We studied how utterance-level (Gricean) expectations about informative communication are integrated with common ground information. Adults’ and children’s learning was best predicted by a model in which both sources of information traded-off flexibly. Alternative models that considered only one source of information made substantially worse predictions.

Cue integration is an integral part of language comprehension and learning [as well as perception more generally; @trommershauser2011sensory]. As such, it has been extensively studied in recent decades. The focus of this work has usually been on how adults combine perceptual, semantic or syntactic information [e.g. @tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line]. We extend the study of linguistic cue integration to pragmatics. Most importantly, however, we present a substantive theory of the integration process itself. Real world language comprehension and learning happens in socially dynamic and complex situations which inevitably require integrating multiple pragmatic information sources. The integration model provides a formal description of the (hypothetical) psychological representations that may underlie information integration. As such, our work complements theorizing about information integration in other domains of language comprehension [e.g. @fourtassi2020optimal; @mcclelland2006there; @smith2017multimodal].

>>>>

abstract: |
  Language is learned in complex social settings where listeners must reconstruct speakers’ intended meanings from context. To navigate this challenge, children can use pragmatic reasoning to learn the meaning of unfamiliar words. One important challenge for pragmatic reasoning is that it requires integrating multiple information sources. Here we study this integration process. We isolate two sources of pragmatic information and -- using a probabilistic model of conversational reasoning -- formalize how they should be combined and how this process might develop. We use this model to generate quantitative predictions, which we test against new behavioral data from three- to five-year-old children (N = 243) and adults (N = 694). Results show close numerical alignment between model predictions and data. Furthermore, the model provided a better explanation of the data compared to simpler alternative models assuming that participnats selectively ignore one information source. This work integrates distinct sets of findings regarding early language and suggests that pragmatic reasoning models can provide a quantitative framework for understanding developmental changes in language learning.

@@

Children learning their first language make inferences about intended meanings based on utterance-level and common ground information -- both for language understanding and language learning [@bohn2019pervasive; @clark2009first; @tomasello2008origins]. Starting very early, infants expect adults to produce utterances in a cooperative way [@behne2005one], and expect language to be carrying information [@vouloumanos2012twelve]. By age two, children are sensitive to the informativeness of communication [@o2001two]. By age three, children can use this expectation to make pragmatic inferences [@stiller2015ad; @yoon2019role] and to infer novel word meanings [@frank2014inferring]. And although older children continue to struggle with some complex pragmatic inferences until age five and beyond [@noveck2001children], an emerging consensus identifies these difficulties as stemming from difficulties reasoning about linguistic alternatives rather than pragmatic deficits [@skordos2016children; @horowitz2018trouble; @barner2011accessing]. Thus, children’s ability to reason about utterance-level pragmatics is present at least by ages three to five, and possibly substantially younger.

Evidence for the use of common ground^[Common ground has traditionally been defined in recursive terms: in order to be part of common ground, some piece of information has to be not just known to both interlocutors but also known to both to be shared between them [@clark1996using]. Numerous studies probed the role of sharedness of information and found that it plays a critical role in communicative interactions [@brown2009partner; @hanna2003effects; @heller2016perspective; @mozuraitis2015privileged]. Based on this literature, one might argue that the term common ground should be restricted to describe situations in which the sharedness aspect is directly tested. However, most of this work is focused on online perspective taking. In this paper, we use the term common ground to refer to shared information that is built up over the course of an interaction - something that is likely easier for children [@matthews2006effect]. We assume that the consequence of a direct interaction (with matching perspectives) between the speaker and the listener is that information is mutually manifest; that is, not just known to both interlocutors but also assumed to be shared between them [@bohn2018common] and hence part of common ground. Thus, since this information is unproblematically in common ground, we can focus on how this information integrates with other pragmatic information sources.] information by young children is even stronger: SPeaker specific contextual expectations guide how infants produce non-verbal gestures and interpret ambiguous utterances [@bohn2018social; @saylor2011s]. For slightly older children, common ground – in the form of knowledge about discourse novelty, preferences, and even discourse expectations – also facilitates word learning [@akhtar1996role; @bohn_le_peloquin_koymen_frank_2020; @saylor2009preschoolers; @sullivan2019discourse].

@@

Within the RSA framework, one way to incorporate common ground is to treat it as a conversational prior. That is, to assume that social interactions result in prior distributions over possible intended meanings which are specific to a particular speaker. Following this logic, a natural locus for information integration within probabilistic models of pragmatic reasoning is the trade-off between the prior probability of a meaning and the informativeness of the utterance. This trade off between contextual factors during word learning is a unique aspect that is not addressed by other computational models of word learning, which have focused on learning from cross-situational, co-occurrence statistics [@fazly2010probabilistic; @frank2009using] or describing generalizations about word meaning [@xu2007word].

@@

We start by replicating previous findings with adults showing that listeners make pragmatic inferences based on non-linguistic properties of utterances in isolation (experiment 1). Then we show that adults make inferences based on common ground information (experiment 2A and 2B). We use data from these experiments as parameters to generate a priori predictions from RSA models about how utterance information and conversational priors should be integrated. We consider three models that make different assumptions about the integration process: In the *integration model*, the two information sources are integrated with one another. The other two models are lesioned models which assume that participants focus on one type of information and disregard the other whenever they are presented together. According to the *no conversational prior* model, participants focus only on the utterance information and in the *no informativeness* model, only the conversational prior is considered. We compare predictions from these models to new empirical data from experiments in which utterance and common ground information are manipulated simultaneously (Experiment 3 and 4).

@@

Taken together, this work makes two primary contributions: first, it shows that both adults and children integrate utterance-level information with conversational priors. Second, it uses Bayesian data analysis within the RSA framework to provide a model for understanding the multiple loci for developmental change in complex behaviors like contextual communication.

@@

We preregistered sample sizes, inferential statistical analysis and computational models for all experiments. All deviations from the registered analysis plan are explicitly mentioned. All analyses were run in `R` [@R-base]. All p-values are based on two sided analysis. Cohen's d (computed via the function `cohensD`) was used as effect size for t-tests. Frequentist logistic GLMMs were fit via the function `glmer` from the package `lme4` [@R-lme4] and had a maximal random effect structure conditional on model convergence. Details about GLMMs including model formulas for each experiment can be found in the Supplementary Material.

All models and model comparisons were implemented in `WebPPL` [@dippl] using the `R` package `rwebppl` [@R-rwebppl]. Probabilistic models were evaluated using Bayesian data analysis [@lee2014bayesian], also implemented in `WebPPL`. In experiment 3, 4 and 7, we compared probabilistic models based on Bayes Factors - the ratio of the marginal likelihoods of each model given the data. Details on models, including information about priors for parameter estimation and Markov chain Monte Carlo settings can be found in the Supplementary Material available online. Code to run the models is available in the associated online repository.

@@

Experiments 1 and 2 confirmed that adults make pragmatic inferences based on information provided by the utterance as well as by common ground and provided quantitative estimates of the strength of these inferences for use in our model. We modeled the integration of utterance informativity and conversational priors (representing common ground) as a process of socially-guided probabilistic inference, using the results of experiments 1 and 2 to inform key parameters of a computational model. The Rational Speech Act (RSA) model architecture introduced by Frank and Goodman [-@frank2012predicting] encodes conversational reasoning through the perspective of a listener (“he” pronoun) who is trying to decide on the intended meaning of the utterance he heard from the speaker (“she” pronoun). The basic idea is that the listener combines his uncertainty about the speaker’s intended meaning - a prior distribution over referents $P(r)$ - with his generative model of how the utterance was produced: a speaker trying to convey information to him. To adapt this model to the word learning context, we enrich this basic architecture with a mechanism for expressing uncertainty about the meanings of words (lexical uncertainty) - a prior distribution over lexica $P(L)$ [@bergen2016pragmatic].

@@

In the above equation, the listener is trying to jointly resolve the speaker’s intended referent $r$ and the meaning of words (thus learning the lexicon $\mathcal{L}$). He does this by imagining what a rational speaker would say, given the referent they are trying to communicate and a lexicon. The speaker is an approximately rational Bayesian actor (with degree of rationality $\alpha$), who produces utterances as a function of their informativity. The space of utterances the speaker could produce depends upon the lexicon $P(u|\mathcal{L})$; simply put, the speaker labels objects with the true labels under a given lexicon L (see Supplementary Material available online for details):

@@

This computational model provides a natural avenue to formalize quantitatively how informativeness and conversational priors trade-off during word learning. As mentioned above, we treat common ground as a conversational prior over meanings, or types of referents, that the speaker might be referring to. That is, we assume that the interactions around the referents in the present context (e.g., preference or novelty; experiment 2A and B) result in a person-specific prior distribution over referents. We use the results from experiment 2 to specify this distribution. The in-the-moment, contextual informativeness of the utterance is captured in the likelihood term, whose value depends on the rationality parameter $\alpha$. Assumptions about rationality may change depending on context and we therefore used the data from experiment 1 to specify $\alpha$ (see Supplementary Material available online for details about these parameters).

@@

As a second step, we compared the model predictions to the data. Participants’ average responses were highly correlated with the  predictions from the *integration model* in each condition (Fig \ref{fig:fig3}b). When comparing models, we found that model fit was considerably better for the *integration model* compared to the *no conversational prior model* (Bayes Factor (BF) = `r format(ex3_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex3_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`), suggesting that participants considered and integrated both sources of information.

Finally, we examined the noise parameter for each model. The estimated proportion of random responses according to the *integration model* was `r ex3_pragm_model_noise_param%>%pull(mean)` (95% Highest Density Interval (HDI): `r ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r ex3_pragm_model_noise_param%>%pull(ci_upper)`). This parameter was substantially lower for the *integration model* compared to the alternative models (*no conversational prior model*: `r ex3_flat_prior_model_noise_param%>%pull(mean)` [`r ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex3_prior_only_model_noise_param%>%pull(mean)` [`r ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex3_prior_only_model_noise_param%>%pull(ci_upper)`]), lending additional support to the conclusion that the *integration model* better captured the behavioral data. Rather than explaining systematic structure in the data, the alternative models achieved their best fit only by assuming a very high level of noise.

@@

To test if the *integration model* makes accurate predictions for different combinations, we first replicated and then extended the results of experiment 3 to a broader range of experimental conditions. Specifically, we manipulated the strength of the common ground information (3 levels - strong, medium and weak - for preference and 2 levels - strong and medium - for novelty) by changing the way the speaker interacted with the objects prior to the request. The procedural details and statistical analysis for these these manipulations are described in the Supplementary Material available online. For experiment 4, we paired each level of prior strength manipulation with the informativeness inference in the same way as in experiment 3. This resulted in a total of 20 conditions, for which we generated a priori model predictions in the same way as in experiment 3. That is, we conducted a separate experiment for each level of prior strength and common ground manipulation to estimate the prior probability of each object following this particular manipulation (analogous to experiment 2). This prior distribution was then passed through the model for the congruent and incongruent conditions, resulting in a unique prediction for each of the 20 conditions. Given the graded nature of the prior manipulations, experiment 4 basically tests how well the model performs with different types of prior distributions.

@@

The direct replication of experiment 3 within experiment 4 showed a very close correspondence between the two rounds of data collection (see Fig \ref{fig:fig3}c). GLMM results for experiment 4 can be found in the Supplementary Material available online. Here we focus on the analysis based on the probabilistic models. Model predictions from the *integration model* were again highly correlated with the average response in each condition (see Fig \ref{fig:fig3}d). We evaluated model fit for the same models as in experiment 3 and found again that the *integration model* fit the data much better compared to the *no conversational prior* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_flat_prior_noise), digits = 2)`) or the *no informativeness model* (BF = `r format(ex4_bf%>%pull(pragmatic_noise_prior_only_noise), digits = 2)`). The inferred level of noise based on the data for the *integration model* was `r ex4_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r ex4_pragm_model_noise_param%>%pull(ci_lower)` - `r ex4_pragm_model_noise_param%>%pull(ci_upper)`), which was similar to experiment 3 and again lower compared to the alternative models (*no conversational prior model*: `r ex4_flat_prior_model_noise_param%>%pull(mean)` [`r ex4_flat_prior_model_noise_param%>%pull(ci_lower)` - `r ex4_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r ex4_prior_only_model_noise_param%>%pull(mean)` [`r ex4_prior_only_model_noise_param%>%pull(ci_lower)` - `r ex4_prior_only_model_noise_param%>%pull(ci_upper)`]).

@@

In experiment 6, we assessed whether children use common ground information to identify the referent of a novel word. We tested children only with the preference manipulation^[We initially tested children with the novelty as well as the preference manipulation. We found that children made the basic inference in that they selected the object that was preferred by or new to the speaker, but found little evidence that children distinguished between requests made by the same speaker or a different speaker in the case of novelty. This finding contrasts with earlier work [@diesendruck2004two]. Since our focus was on how children integrate informativeness and conversational priors resulting from common ground, we did not follow up on this finding but dropped the novelty manipulation and focused on preference for the remainder of the study.]. The procedure for children was identical to the preference manipulation for adults. Children received eight trials, four with the same and four with a different speaker.

@@

child_ex3_bf <- readRDS("../../stats/saves/child_ex3_model_comparison.rds") %>%
  group_by(parameter)%>%
  select(-ci_lower, -ci_upper)%>%
  spread(model, logP) %>%
  mutate("pragmatic_flat_prior" = exp(pragmatic - flat_prior),
         "pragmatic_prior_only" = exp(pragmatic - prior_only),
         "flat_prior_prior_only" = exp(flat_prior - prior_only)) %>%
  select(-pragmatic,-flat_prior, -prior_only)

@@

Analyses comparing the model predictions from the probabilistic models to the data suggest that children flexibly integrate conversational priors and informativity information. Furthermore, this integration process is accurately captured by the *integration model* at least for four-year-olds. For the correlational analysis, we binned model predictions and data by year. There was a substantial correlation between the predicted and measured average response for four-year-olds, but less so for three-year-olds (Fig \ref{fig:fig5}b). One of the reasons for the latter was the low variation between conditions. For the model comparison, we treated age continuously. As with adults, we found a much better model fit for the *integration model* compared to the *no conversational prior* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_flat_prior), digits = 2)`) or the *no informativeness model* (BF = `r format(child_ex3_bf%>%filter (parameter=="noise")%>%pull(pragmatic_prior_only), digits = 2)`).

The inferred level of noise based on the data for the integration model was `r child_ex3_pragm_model_noise_param%>%pull(mean)` (95% HDI: `r child_ex3_pragm_model_noise_param%>%pull(ci_lower)` - `r child_ex3_pragm_model_noise_param%>%pull(ci_upper)`), which was lower compared to the alternative models considered (*no conversational prior model*: `r child_ex3_flat_prior_model_noise_param%>%pull(mean)` [`r child_ex3_flat_prior_model_noise_param%>%pull(ci_lower)` - `r child_ex3_flat_prior_model_noise_param%>%pull(ci_upper)`]; *no informativeness model*: `r child_ex3_prior_only_model_noise_param%>%pull(mean)` [`r child_ex3_prior_only_model_noise_param%>%pull(ci_lower)` - `r child_ex3_prior_only_model_noise_param%>%pull(ci_upper)`]) but numerically higher than that of adults (see Fig \ref{fig:fig5}c).

The high level of inferred noise moved the model predictions for children in all conditions close to chance level. We therefore compared two additional sets of models with different parameterizations of the noise parameter that emphasized differences between conditions in the model predictions more (see Supplementary Material and Fig \ref{fig:fig5}a). This analysis was not pre-registered. Parameter free models did not include a noise parameter and developmental noise models allowed the noise parameter to change with age. In each case, the *integration model* provided a better fit compared to the alternative models (*no conversational prior*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_flat_prior), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_flat_prior), digits = 2)`; *no informativeness*: parameter free BF = `r format(child_ex3_bf%>%filter (parameter=="parameter free")%>%pull(pragmatic_prior_only), digits = 2)`, developmental noise BF = `r format(child_ex3_bf%>%filter (parameter=="developmental noise")%>%pull(pragmatic_prior_only), digits = 2)`). The developmental noise parameter for the integration model decreased with age, suggesting that older children behaved more in line with model predictions compared to younger children (see Fig \ref{fig:fig5}d).

```{r fig5, include = T, fig.align = "center", fig.cap = "Results from experiment 7 for children. (a) Model predictions (with 95\\% HDIs) and data across age in the four conditions. Transparent black dots show data from individual participants and black lines show conditional means of the data with 95\\% CI. Black diamonds show the mean of the data for age bins by year and error bars show 95\\% CIs. (b) Correlation between model predictions (with noise parameter) and condition means binned by year (with 95\\% HDIs). For 4-year-olds, two conditions yielded the same data means and model predicitons and are thus plotted on top of each other. (c) Posterior distribution of the noise parameter for the different models. (d) Developmental trajectory of the noise paramter for the three developmental noise models; trajectories are based on MAPs of the posterior distribution for the intercept and slope.", out.width = "100%"}
knitr::include_graphics("./figures/fig5.png")
```

@@

Integrating multiple sources of information is an integral part of human communication. To infer the intended meaning of an utterance, listeners must combine their knowledge of communicative conventions (semantics and syntax) with social expectations about their interlocutor. This integration is especially vital in early language learning, when the different varieties of pragmatic information are among the most important sources [@bohn2019pervasive]. But how are pragmatic cues integrated during word learning? Here we used a Bayesian cognitive model to formalize this integration process. We studied how utterance-level (Gricean) expectations about informative communication are integrated with conversational priors (resulting from common ground). Adults’ and children’s learning was best predicted by a model in which both sources of information traded-off flexibly. Alternative models that considered only one source of information made substantially worse predictions.

Cue integration is an integral part of language comprehension and learning [as well as perception more generally; @trommershauser2011sensory]. As such, it has been extensively studied in recent decades. The focus of this work has usually been on how adults combine perceptual, semantic or syntactic information [e.g. @tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line]. We extend the study of linguistic cue integration to pragmatics. Most importantly, however, we present a substantive theory of the integration process itself. Real world language comprehension and learning happens in socially dynamic and complex situations which inevitably require integrating multiple pragmatic information sources. The *integration model* provides a formal description of the (hypothetical) psychological representations that may underlie information integration. As such, our work complements theorizing about information integration in other domains of language comprehension [e.g. @fourtassi2020optimal; @mcclelland2006there; @smith2017multimodal].
